---
title: "Data Cleaning"
output: html_document
date: "2022-12-05"
---

```{r setup, include=FALSE}
library(tidyverse)
library(tidyselect)
library(caTools)

#Reminder:
#probation_cat are potentially dangerous categories that can be removed later
```

Try to reduce the total number of variables.
```{r}
surgery_df <- read.csv("Dataset.csv")

surgery_df_1 <- surgery_df%>%
  select(-c(encounter_id,
            hospital_admit_source,          #Redundant with icu_admit_source
            icu_id, 
            icu_stay_type,                  #Same type of variable as readmission status. However, upon survival, subject should still count as population
            readmission_status,             
            contains("_apache"),            #apache covariates removed since those are used in another pre-existing model to estimate survivability rate
            contains("invasive"),           #non-invasive methods are typically standard, but invasive methods are used when required. Using combined col.
            apache_2_diagnosis,             #apache diagnosis score removed because the numbers in the data don't fit in the tool range for some reason.
            apache_3j_diagnosis))

df_minmax <-surgery_df_1%>%           #I want the averages. Don't particularly need min max for the model.
  select(contains("min"),
         contains("max"))

df_minmax <- lapply(df_minmax, as.numeric)
df_minmax <- as.data.frame(df_minmax)


longer_minmax <- pivot_longer(df_minmax, cols = contains("_"), 
  names_pattern = "(.*)_(min|max)",
  names_to = c("dimension", ".value"))  

longer_minmax <- longer_minmax %>%
  mutate(avg = (min + max)/2)

df_avg <- longer_minmax %>%
  select(-min,-max)%>%
  group_by(dimension) %>%
  mutate(row = row_number()) %>%
  pivot_wider(names_from = dimension, values_from = avg,names_glue="{dimension}_{.value}") %>%
  select(-row)

surgery_df_avgs <- cbind(surgery_df_1, df_avg)
surgery_df_avgs <- surgery_df_avgs %>%
  select(-c(contains("max"),contains("min")))

```
Count NAs

A significant amount of missing data appears to be in the categories that require lab tests. It is most likely that the tests simply weren't
performed for these individuals. I could get rid of them, but I'd rather keep whatever data I have if possible.
To keep as much data as possible, I am going to impute the mean average as a replacement for NAs. I will also make a list of all variables with more than
20k NAs to remove later if need be. I will also have to remove rows that have NAs in categorical variables since those are not imputable.
```{r}
surg_na <- sapply(surgery_df_avgs, function(count_na) sum(is.na(count_na)))
probation_cat <- names(surg_na)[surg_na > 20000]

surg_na

surgery_df_x <- surgery_df_avgs %>%
  filter(!is.na(cirrhosis))%>%
  mutate_if(is.numeric, funs(replace(., is.na(.), median(., na.rm = TRUE))))   #Imputed by median instead of mean in case of huge outliers

surg_na <- sapply(surgery_df_x, function(count_na) sum(is.na(count_na))) #check to see if all missing values are gone

write.csv(surgery_df_x, file = "cleaned_surgery.csv")
```




